“We propose a GenAI‑assisted configuration and pattern generator for the existing IRD Tkinter email extractor. When the extractor encounters new broker formats or fails in UAT, a RAG pipeline will retrieve similar historical emails and current config patterns, and an LLM will propose new JSON/regex extraction rules and explain their behavior. These suggestions are reviewed and approved by engineers before deployment; the LLM never books trades or runs in the production extraction path. The goal is to reduce manual pattern authoring effort, speed up onboarding of new templates and products, and make the rulebase more maintainable, while keeping all risk‑sensitive decisions and trade booking logic in the existing deterministic engine.”



OFF-MARKET RATE CHECKING INTELLIGENCE (OMRCI)
Multi-Asset Hybrid Risk Control System (IRD, GFX, GBM)
AI Governance Use Case Submission

VERSION 2.0 | NOVEMBER 28, 2025
CLASSIFICATION: INVESTMENT BANKING | INTERNAL RISK CONTROL
STATUS: POC COMPLETE (CELS) | UAT IN PROGRESS (IRD/Tkinter)

═══════════════════════════════════════════════════════════════════════════════

EXECUTIVE SUMMARY

PROPOSAL TITLE
Off-Market Rate Checking Intelligence (OMRCI): Multi-Asset Risk Validation Ecosystem

USE CASE IN ONE SENTENCE
A unified risk control platform that orchestrates product-specific validation logic: GenAI-powered email reconstruction for complex IRD negotiations (via Tkinter) and Semantic Justification Analysis for GFX/GBM high-volume flows (via CELS).

CURRENT STATUS
• Custom Exception Linkage System (CELS): POC Complete. Proven capability to link exceptions to historical patterns.
• IRD Email Extraction (Tkinter): Currently in UAT. Validating email-based rate justifications.
• GenAI Intelligence Layer: PROPOSED. The final component to automate justification analysis across both flows.

STRATEGIC ALIGNMENT
✓ Critical Risk Control: Prevents "off-market" booking errors and regulatory breaches.
✓ Product-Specific Fit: Acknowledges that IRD (Complex/Email-based) and GFX (High-vol/Source-based) require different validation methods.
✓ Asset Utilization: Leverages the CELS engine (already built) and Tkinter App (in UAT), rejecting the costly and ill-fitted SWB alternative.

═══════════════════════════════════════════════════════════════════════════════

SECTION 1: THE DUAL-FLOW PROBLEM

1.1 THE ASSET CLASS DIVERGENCE
Generic tools like SWB assume all trades are documents. In reality, Off-Market Rate checking requires two distinct workflows:

FLOW A: IRD (Interest Rate Derivatives - Swaps, Swaptions)
• Nature: Low volume, High complexity.
• Data Source: Unstructured Email Conversations.
• Validation Method: "Conversational Reconstruction."
• Risk: Terms negotiated over days; justification buried in thread (e.g., "Discount agreed due to CSA terms").

FLOW B: GFX & GBM (Global FX, Global Business Markets)
• Nature: High volume, Standardized.
• Data Source: Trader Notes + External Market Sources (Bloomberg/Reuters).
• Validation Method: "Source Corroboration."
• Risk: Trader enters 1-line note (e.g., "Distressed liquidity"). System must cross-reference this against market volatility snapshots.

1.2 WHY SWB FAILS FOR BOTH
• Fails IRD: SWB extracts fields from single docs. It cannot reconstruct a 5-email negotiation thread to find the *reason* for a price deviation.
• Fails GFX/GBM: SWB processes documents. It cannot connect to live Bloomberg market data to verify if a "Liquidity" justification is valid.

═══════════════════════════════════════════════════════════════════════════════

SECTION 2: PROPOSED SOLUTION - THE OMRCI ECOSYSTEM

We propose a centralized intelligence layer that sits on top of our two existing ingestion channels.

2.1 THE CORE ENGINE: CELS (Status: POC Complete)
The "Brain" of the operation.
• Ingests exceptions from risk engines.
• Links current exceptions to historical approvals (e.g., "Client X standing approval").
• Holds the "State" of the investigation.

2.2 INPUT CHANNEL A: IRD / Tkinter (Status: UAT In-Progress)
The "Reader" for complex deals.
• Workflow: Analyst opens Tkinter app → Fetches email thread for Trade ID.
• GenAI Task: "Read this email chain. Extract the justification for the 5bps deviation. Identify the approver."
• Output: Structured justification sent to CELS.

2.3 INPUT CHANNEL B: GFX / GBM (Status: Standard Integration)
The "Validator" for flow trading.
• Workflow: System pulls Trader Note + Market Data Snapshot.
• GenAI Task: "Trader justification is 'Volatile Market'. Compare against VIX/Volatility index at trade time. Is this justification plausible?"
• Output: Validation Score (Green/Red) sent to CELS.

═══════════════════════════════════════════════════════════════════════════════

SECTION 3: DETAILED GENAI USE CASE

3.1 USE CASE 1: CONVERSATIONAL REASONING (IRD)
• Input: 5-email thread extracted by Tkinter (currently in UAT).
• Challenge: Determining *which* email contains the final agreed price and *why* it deviates.
• GenAI Role:
  - Summarize the negotiation timeline.
  - Pinpoint the specific sentence justifying the rate (e.g., "Agreed to mid-market minus 2 due to early termination").
  - Map this unstructured text to a Risk Category (e.g., "Commercial Agreement").

3.2 USE CASE 2: SEMANTIC VALIDATION (GFX/GBM)
• Input: Short text Trader Note (e.g., "Leg 2 off mkt/package") + Market Context.
• Challenge: High volume validation.
• GenAI Role:
  - Classify the note.
  - Cross-reference: If note says "Package Trade", check if Strategy ID links to a multi-leg deal.
  - Cross-reference: If note says "Market Move", check Reuters feed for price spikes at that minute.

═══════════════════════════════════════════════════════════════════════════════

SECTION 4: PROJECT STATUS & TIMELINE

PHASE 1: FOUNDATION (COMPLETE)
✓ CELS Core Logic: Developed & Validated (POC Done).
✓ Exception Linkage Logic: 95% Accuracy in testing.

PHASE 2: IRD INTEGRATION (IN PROGRESS)
► Tkinter App Development: Complete.
► UAT Testing: 60% Complete.
► GenAI Integration: PENDING APPROVAL (This Request).

PHASE 3: GFX/GBM EXPANSION (PLANNED)
• Connect CELS to Market Data API.
• Deploy GenAI Semantic Validator for high-volume notes.

═══════════════════════════════════════════════════════════════════════════════

SECTION 5: GOVERNANCE & CONTROL

5.1 HUMAN-IN-THE-LOOP
• IRD: GenAI suggests the justification from the email. Analyst MUST click "Confirm" in the Tkinter UI.
• GFX/GBM: GenAI flags "Implausible Justifications" (e.g., Trader claims volatility, but market was flat). Analyst investigates flags.

5.2 AUDITABILITY
• CELS stores the "Golden Record."
• Every AI suggestion is logged alongside the specific email ID or Market Data snapshot used to generate it.

5.3 DATA SOVEREIGNTY
• Email data processed on-prem / private cloud.
• No sensitive client data used for model training.

═══════════════════════════════════════════════════════════════════════════════

SECTION 6: VALUE PROPOSITION

WHY THIS SPECIFIC APPROACH?
1. Agility: We are already in UAT with IRD/Tkinter. Switching to SWB now would scrap 6 months of work and fail to meet the specific "Conversational" requirement of IRD.
2. Coverage: SWB cannot handle the "Source-based" validation required for GFX/GBM. Only our CELS + GenAI approach covers the full Investment Bank portfolio.
3. Efficiency: Automates the complex reading task (IRD) and the high-volume validation task (GFX), saving estimated $1.5M/year in analyst time.

RECOMMENDATION:
Approve GenAI integration for the OMRCI ecosystem to support the active UAT of IRD/Tkinter and the Phase 3 rollout for GFX/GBM.
═══════════════════════════════════════════════════════════════════════════════
