OFF-MARKET RATE CHECKING INTELLIGENCE (OMRCI)
Hybrid Neuro-Symbolic AI for Investment Banking Risk Control
AI Governance Use Case Submission

VERSION 1.0 | NOVEMBER 28, 2025
CLASSIFICATION: INVESTMENT BANKING | INTERNAL RISK CONTROL

═══════════════════════════════════════════════════════════════════════════════

EXECUTIVE SUMMARY

PROPOSAL TITLE
Off-Market Rate Checking Intelligence (OMRCI): Hybrid AI for Trade Rate Validation & Exception Management

USE CASE IN ONE SENTENCE
A specialized risk control system combining existing Off-Market Rate (OMR) exception linkage with GenAI-powered trade reconstruction to validate complex rate deviations that generic extraction tools (SWB) cannot process.

KEY METRICS (Year 1 Projected)
• False Positive Rate Reduction: 35% → 5%
• Analyst Investigation Time: 15 minutes → 3 minutes per trade
• Risk Coverage: 100% of off-market trades validated against email context
• Annual Cost Savings: $1.5M (Risk Analyst Efficiency)
• Regulatory Compliance: Enhanced audit trail for MRA (Matter Requiring Attention) resolution

STRATEGIC ALIGNMENT
✓ Critical Risk Control (prevents significant financial loss)
✓ Regulatory Requirement (Market Abuse Regulation / Dodd-Frank)
✓ Leverages Custom Exception Logic (cannot be replaced by generic SWB)
✓ Cost-Efficient (GenAI used only for contextual reconstruction)

═══════════════════════════════════════════════════════════════════════════════

SECTION 1: PROBLEM STATEMENT & SWB LIMITATION

1.1 THE "OFF-MARKET" CHALLENGE
Investment Banking traders often execute trades at rates that deviate from the standard market price due to valid reasons (e.g., structured deals, package trades, distress sales). However, these trades flag as "Off-Market" in risk systems.

The Off-Market Rate Checking Team must validate:
1. Is the rate actually off-market?
2. Is there a valid reason documented in the email negotiation?
3. Was the deviation approved by a supervisor?

1.2 WHY SWB CANNOT HANDLE THIS
SWB (Strategic Workbench / Smart Workbench) is a generic document extraction tool designed to turn PDFs into rows and columns. It fails in this domain because:

FAILURE POINT 1: NO "RATE LOGIC"
SWB extracts "Price: 99.5". It does not know if 99.5 is off-market. It cannot compare 99.5 against the Bloomberg mid-rate at 10:42 AM on the trade date.
Our existing system DOES this.

FAILURE POINT 2: CONTEXT BLINDNESS
Valid off-market trades are justified in conversational email text:
"Agreed to 50bps discount due to illiquidity of the underlying asset."
SWB extracts fields; it does not extract *justifications* or *reasoning*.

FAILURE POINT 3: EXCEPTION LINKAGE
We already have a system that links current exceptions to historical patterns (e.g., "Client X always trades off-market on distressed assets").
SWB is stateless; it treats every trade as new. Replacing our system with SWB would destroy this institutional memory.

═══════════════════════════════════════════════════════════════════════════════

SECTION 2: EXISTING ASSETS (THE "DEFENSE")

2.1 CUSTOM EXCEPTION LINKAGE SYSTEM (CELS)
We have built a specialized application for the Off-Market Checking Team that:
• Ingests trade data + market data (Bloomberg/Reuters)
• Calculates deviation (Z-score) automatically
• Links new exceptions to historical approvals (e.g., "This bond has a standing approval for 2% deviation")
• **Status:** PRODUCTION
• **Value:** Saves 10,000 hours/year in manual checking

2.2 TKINTER EXTRACTION APP
A lightweight tool used by analysts to pull specific email threads related to flagged trades.
• **Status:** PRODUCTION
• **Value:** Bridges the gap between the risk system and Outlook.

═══════════════════════════════════════════════════════════════════════════════

SECTION 3: PROPOSED SOLUTION (OMRCI)

We propose ENHANCING our existing CELS + Tkinter ecosystem with a targeted GenAI module. We are NOT building a generic extraction tool; we are building a "Justification Analyzer."

ARCHITECTURE:
1. **Trade Flagged:** CELS identifies a trade 3% off-market.
2. **Context Retrieval:** Tkinter app pulls the relevant email thread.
3. **GenAI Analysis (The New Part):**
   • Prompt: "This trade is 3% below market. Read the email chain. Is there a documented justification for this price deviation? Extract the reason and the approver."
4. **Validation:** Analyst reviews the GenAI-extracted justification alongside the market data.

WHY THIS IS BETTER THAN SWB:
• **SWB:** Extracts "Price: 98.0". Result: Exception remains open. Analyst must read email manually.
• **OMRCI:** Extracts "Reason: Distressed Sale approved by MD". Result: Exception auto-resolved (pending 1-click review).

═══════════════════════════════════════════════════════════════════════════════

SECTION 4: FINANCIAL JUSTIFICATION

INVESTMENT: $120,000 (GenAI Integration)
SAVINGS: $1.5M / Year

Breakdown:
• **Analyst Efficiency:** 50 analysts × 20% time saved = $1M
• **Risk Avoidance:** Prevention of 1 major regulatory fine ($500k+ value)
• **SWB License Avoidance:** Avoids $200k/year in unnecessary vendor fees for a tool that doesn't meet requirements.

═══════════════════════════════════════════════════════════════════════════════

SECTION 5: GOVERNANCE & COMPLIANCE

DATA PRIVACY:
• No client PII sent to LLM (masked).
• Only trade economics and justification text processed.

MODEL RISK:
• GenAI provides *suggestions* only.
• Risk Analyst retains 100% decision authority.
• Audit trail logs exactly what the AI saw and what the human decided.

REGULATORY ALIGNMENT:
• Meets "Supervisory Control" requirements by ensuring every off-market trade has a documented, validated reason.

═══════════════════════════════════════════════════════════════════════════════

CONCLUSION

Forcing SWB on the Off-Market Rate Checking Team would be a regression. SWB cannot perform market data lookups, deviation calculations, or justification extraction.

Our proposal leverages the firm's existing investment in the Custom Exception Linkage System and adds a surgical GenAI layer to handle the one thing computers struggle with: reading human justifications. This is the optimal path for risk reduction and cost efficiency.

RECOMMENDATION: APPROVE OMRCI USE CASE.
═══════════════════════════════════════════════════════════════════════════════